# **Reproducible Research: Peer Assessment 1**

## Loading and preprocessing the data
Load the data and Process/transform the data (if necessary) into a format suitable for your analysis.
```{r}
unzip(zipfile='activity.zip') 
dt <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day, Make a histogram of the total number of steps taken each day
```{r meanplot}
stepsum <- with(dt, tapply(steps,date,FUN=sum,na.rm=TRUE))
hist(stepsum, breaks=20, xlab='Total Steps per Day', main='Total Number of Steps Taken Each Day')
```
Calculate and report the mean and median of the total number of steps taken per day.
```{r meanmedian}
mean(stepsum,na.rm=TRUE)
median(stepsum,na.rm=TRUE)
```
## What is the average daily activity pattern?
Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis).
```{r}
avestep <- aggregate(x=dt$steps,by=list(dt$interval),FUN=mean,na.rm=TRUE)
with(avestep, plot(Group.1,x,type='l',xlab='Interval',ylab='Average Steps Taken',main='AVerage Steps Taken per 5 Minute Interval'))
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxsteps}
avestep[which.max(avestep$x),]
```
## Imputing missing values
Calculate and report the total number of missing values in the dataset.
```{r}
sum(is.na(dt))
```
Create a new dataset that is equal to the original dataset but with the missing data filled in with the mean value.
```{r dplyr,echo=FALSE,results='hide',message=FALSE}
library(dplyr)
```

```{r completetable}
meanreplace <- function(x) 
  replace(x, is.na(x), mean(x, na.rm = TRUE))
avestep.new<- dt %>% group_by(interval) %>% mutate(steps= meanreplace(steps))
head(avestep.new)
```
Make a histogram of the total number of steps taken each day.
```{r newmeanplot}
stepsum.new <- with(avestep.new, tapply(steps,date,FUN=sum,na.rm=TRUE))
hist(stepsum.new, breaks=20, xlab='Total Steps per Day', main='Total Number of Steps Taken Each Day (Complete Data)')
```
Calculate and report the mean and median total number of steps taken per day.
```{r newmeanmedian}
mean(stepsum.new,na.rm=TRUE)
median(stepsum.new,na.rm=TRUE)
```
After the mean is added to the NA values, it was seen that the mean and median increased to **10766.19**.

## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r weekday}
dt$date <- as.Date(dt$date)
dt$week <- ifelse(weekdays(dt$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
```
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r ggplot2,echo=FALSE,results='hide',message=FALSE}
library(ggplot2)
```

```{r}
avestepweek <- aggregate(x=dt$steps,by=list(dt$interval,dt$week),FUN=mean,na.rm=TRUE)
ggplot(avestepweek,aes(x=Group.1,y=x,color=Group.2))+geom_line()+facet_grid(Group.2 ~.)+labs(x='Interval',y='Average Steps Taken',title='Average Steps Taken per 5 Minute Interval in Weekend/Weekday',col='Weekday')